Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-01-13 paper.qmd
Added what is meant by single cell data
Added number of gene experessions for PBMC3k

XXX What is single cell data????

Clustering of single-cell data is used to identify groups of cells with similar expression profiles. NLDR often used to summarise the discovered clusters, and help to understand the results. The purpose of this example is to *illustrate how to use our method to help decide on an appropriate NLDR layout that accurately represents the data*. The cluster results of Human Peripheral Blood Mononuclear Cells (PBMC3k) in @chen2023 are examined. There are $2622$ single cells, with XXX gene expressions. Following their pre-processing, UMAP was performed using nine principal components. @fig-NLDR-variety (a) is the reproduction of the published plot, with colour indicating the clustering results overlaid on the NLDR layout. The question is whether this accurately represents the cluster structure in the data. Our method can help here, and also help to provide a better \gD{} layout, as needed.

>>>>

<!-- XXX What is single cell data???? -->

Single-cell data refers to the characteristics of individual cells within a population (@haque2017). These data also provides insights at the level of individual cells, enabling deeper understanding of gene expression.

Clustering of single-cell data is used to identify groups of cells with similar expression profiles. NLDR often used to summarise the discovered clusters, and help to understand the results. The purpose of this example is to *illustrate how to use our method to help decide on an appropriate NLDR layout that accurately represents the data*. The cluster results of Human Peripheral Blood Mononuclear Cells (PBMC3k) in @chen2023 are examined. There are $2622$ single cells, with $1000$ gene expressions. Following their pre-processing, UMAP was performed using nine principal components. @fig-NLDR-variety (a) is the reproduction of the published plot, with colour indicating the clustering results overlaid on the NLDR layout. The question is whether this accurately represents the cluster structure in the data. Our method can help here, and also help to provide a better \gD{} layout, as needed.
